Survivable Set Connectivity

### 2015-05-13 draft.tex
changed typos

We provide intuition behind the definition of load. Note that in Figure 1, $load(x_1) = 0$, since $\mathrm{map}^{-1}(x_1) = \emptyset$. As an immediate consequence of this, we can see that removing an edge with load 0 does not affect the connectivity of the graph. At a high level, the load measures how necessary an edge is for connectivity. 
\noindent \textbf{Additional Details.} For the remainder of this section, we provide a number of statements regarding R\"{a}cke's trees that will be applied to complete the analysis of the algorithm in Section 6. They are provided here for the sake of completeness, but the reader may skip this section and refer to it if interested while reading the proofs in Section 6.

@@

Recall that $\delta(S)$ is the set of edges spanning the cut $(S,V_T \setminus S)$. This program is almost identical to the program for Survivable Set Connectivity, but with $k$ repalced with 1. The argument for why this program works is the same as before; we are minimizing the total cost of the included edges, and we are ensuring with the constraints that there is a path from $r$ to $t_i$ for some $t_i \in T_i$.

@@

Recall that SSC-LP is the linear programming relaxation of the ILP for SSC, presented in Section 2. Also, $\mathcal{D}$ is the distribution over all possible R\"{a}cke's tree embeddings for a graph $G$, which we can sample from in polynomial time. In the algorithm, we use the variable $\tau$ set to be $\log(h)+k$, $\tau'$, set to be $O(\log n)$, and $q$ as an index.

The algorithm is as follows.

@@

The algorithm first solves the linear programming relaxation, which produces a value of $x_e \in [0,1]$ for each edge $e \in E$. The first step is to zero out all values of $x_e < \frac{k}{n^2}$. After this, the values are scaled down from $x$ to $x'$ by a $\frac{4}{k}$ factor. Since the linear progrom ensures a fractional path of total capacity $k$ from $S_i$ to $T_i$ for each $i$, this $4/k$ scaling ensures a fractional paths of total capacity $4$. Since we argued that a flow in the original graph $G$ corresponds to a flow of the same value in any corresponding R\"{a}cke's tree, the resulting tree embeddings will also have flows of value 4.

@@

The purpose of these scaling constants is to ensure that the RoundGKR procedure, which looks for paths of value 1, will actually see these paths. We preserve the fact that paths with low capacity are chosen with lower probability. The reason why we have $\tau' = O(\log n)$ iterations is because is because  Lemma~\ref{lem:findpath} shows that RoundGKR finds paths in trees which avoid specific edges with probability $\Omega(\frac{1}{\log n})$. $O(\log n)$ repetitions ensure a constant probability of finding such paths.t

@@

Step 1 is a trivial set-up step, and step 2 was Lemma 6~\ref{lem:notbad}. Here, we give the details of step 5.

@@

It seems that there is much work for improving the algorithm, which currently has a $\log^4(n)$ term in the approximation ratio. In an optional section following this conclusion, we give some of our own ideas for potential improvements. We were able to shave one of the $\log(n)$ factors in a very special case, but we did not think that our improvements were substantial enough to deserve a large section in this paper.

@@

If we know $\sum_{i=1}^h |S_i| = S$ is small (the number of sources is small). We can run the algorithm without looking for the nodes where the flow paths turn. In particular, we could treat it as many different instances of the Group Steiner Tree problem. The algorithm would be:
\begin{codebox}
\li Set $H \leftarrow \emptyset$
\li Compute linear program SSC-LP and set to zero any $x_e < \dfrac{k}{n^2}$.
\li Let $x' \leftarrow x*\frac{4}{k}$. 
\li Let $\mathcal{D}$ be the distribution of R\"{a}cke's tree embeddings for $G$ with $x'$.
\li \For $\tau$ rounds \Do
\li Sample $(T, y, \text{map})$ from $\mathcal{D}$. 
\li \For $q = 0, ..., 2\log n$ \Do
\end{codebox}
\begin{codebox*}
\li \For \For $i \in [h]$ and $s \in S_i'$ \Do
\li With probability $\min\{1, \dfrac{4}{2^q}\}$, $q$-mark $s$
\li \If $s$ is $q$-marked \Then
\li Set $y^{s, q}\leftarrow 2^{q+1}y$ and $E^{s,q} \leftarrow \emptyset$.
\li \For $\tau'$ iterations, \Do
\li $E^{s,q} \leftarrow E^{s,q} \cup$ RoundGKR$(T_s, s, y^{s,q})$ \End
\li $H \leftarrow H \cup \text{map}(E^{s,q})$ \End \End \End \End
\li return $H$
\end{codebox*}
The main difference lies in line 8. Instead of iterating through all $v \in V_T$ to find where the flow path turns, we take advantage of knowing where the flow begins. We know all flow paths go from $S_i'$ to $T_i'$, so we make $s \in S_i'$ the root of the tree. It is unnecessary to look for where the flow turns. We still have to iterate through $q= 0, ..., 2\log n$, since the flow of at least $1$ between $S_i'$ and $T_i'$ might be divided among different flow paths starting at different nodes in $S_i'$. This gets rid of the $O(\log n)$ factor in the cost because of the height, and adds a factor of $S$.

>>>>

We provide intuition behind the definition of load. Note that in Figure 1, $load(x_1) = 0$, since $\mathrm{map}^{-1}(x_1) = \emptyset$. As an immediate consequence of this, we can see that removing an edge with load 0 does not affect the connectivity of the graph. At a high level, the load measures how necessary an edge is for connectivity. \\
\noindent \textbf{Additional Details.} For the remainder of this section, we provide a number of statements regarding R\"{a}cke's trees that will be applied to complete the analysis of the algorithm in Section 6. They are provided here for the sake of completeness, but the reader may skip this section and refer to it if interested while reading the proofs in Section 6.

@@

Recall that $\delta(S)$ is the set of edges spanning the cut $(S,V_T \setminus S)$. This program is almost identical to the program for Survivable Set Connectivity, but with $k$ replaced with 1. The argument for why this program works is the same as before; we are minimizing the total cost of the included edges, and we are ensuring with the constraints that there is a path from $r$ to $t_i$ for some $t_i \in T_i$.

@@

Recall that SSC-LP is the linear programming relaxation of the ILP for SSC, presented in Section 2. Also, $\mathcal{D}$ is the distribution over all possible R\"{a}cke's tree embeddings for a graph $G$, which we can sample from in polynomial time. In the algorithm, we use the variable $\tau$ set to be $\log(h)+k$, $\tau'$, set to be $O(\log n)$, and $q$ as an index. The algorithm is as follows.

@@

The algorithm first solves the linear programming relaxation, which produces a value of $x_e \in [0,1]$ for each edge $e \in E$. The first step is to zero out all values of $x_e < \frac{k}{n^2}$. After this, the values are scaled down from $x$ to $x'$ by a $\frac{4}{k}$ factor. Since the linear progrom ensures a fractional path of total capacity $k$ from $S_i$ to $T_i$ for each $i$, this $4/k$ scaling ensures fractional paths of total capacity $4$. Since we argued that a flow in the original graph $G$ corresponds to a flow of the same value in any corresponding R\"{a}cke's tree, the resulting tree embeddings will also have flows of value 4.

@@

The purpose of these scaling constants is to ensure that the RoundGKR procedure, which looks for paths of value 1, will actually see these paths. We preserve the fact that paths with low capacity are chosen with lower probability. The reason why we have $\tau' = O(\log n)$ iterations is because is because  Lemma~\ref{lem:findpath} shows that RoundGKR finds paths in trees which avoid specific edges with probability $\Omega(\frac{1}{\log n})$. $O(\log n)$ repetitions ensure a constant probability of finding such paths.

@@

Step 1 is a trivial set-up step, and step 2 was Lemma~\ref{lem:notbad}. Here, we give the details of step 5.

@@

It seems that there is much room for improvements to the algorithm, which currently has a $\log^4(n)$ term in the approximation ratio. In an optional section following this conclusion, we give some of our own ideas for potential improvements. We were able to shave one of the $\log(n)$ factors in a very special case, but we did not think that our improvements were substantial enough to deserve a large section in this paper.

@@

If we know $\sum_{i=1}^h |S_i| = S$ is small (the number of sources is small). We can run the algorithm without looking for the nodes where the flow paths turn. In particular, we could treat it as many different instances of the Group Steiner Tree problem. The algorithm would be:
\begin{codebox}
\li Set $H \leftarrow \emptyset$
\li Compute linear program SSC-LP and set to zero any $x_e < \dfrac{k}{n^2}$.
\li Let $x' \leftarrow x*\frac{4}{k}$. 
\li Let $\mathcal{D}$ be the distribution of R\"{a}cke's tree embeddings for $G$ with $x'$.
\li \For $\tau$ rounds \Do
\li Sample $(T, y, \text{map})$ from $\mathcal{D}$. 
\li \For $q = 0, ..., 2\log n$ \Do
\li \For \For $i \in [h]$ and $s \in S_i'$ \Do
\li With probability $\min\{1, \dfrac{4}{2^q}\}$, $q$-mark $s$
\li \If $s$ is $q$-marked \Then
\li Set $y^{s, q}\leftarrow 2^{q+1}y$ and $E^{s,q} \leftarrow \emptyset$.
\li \For $\tau'$ iterations, \Do
\li $E^{s,q} \leftarrow E^{s,q} \cup$ RoundGKR$(T_s, s, y^{s,q})$ \End
\li $H \leftarrow H \cup \text{map}(E^{s,q})$ \End \End \End \End
\li return $H$
\end{codebox}
The main difference lies in line 8. Instead of iterating through all $v \in V_T$ to find where the flow path turns, we take advantage of knowing where the flow begins. We know all flow paths go from $S_i'$ to $T_i'$, so we make $s \in S_i'$ the root of the tree. It is unnecessary to look for where the flow turns. We still have to iterate through $q= 0, ..., 2\log n$, since the flow of at least $1$ between $S_i'$ and $T_i'$ might be divided among different flow paths starting at different nodes in $S_i'$. This gets rid of the $O(\log n)$ factor in the cost because of the height, and adds a factor of $S$.
